---
title: "EDA"
output:
  html_document:
    df_print: paged
    css: styles.css
---

```{r setup, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

load_reqs <- function(reqs) {
  for(pkg in reqs) {
    if (!(pkg %in% installed.packages())) { install.packages(pkg) }
    suppressPackageStartupMessages(library(pkg, character.only = T))
  }
}

# ADD LIBRARIES HERE
load_reqs(c("tidyr", "dplyr", "data.table", "purrr", "ggplot2", "corrplot", "gridExtra", "grid", "cowplot", "PerformanceAnalytics", "dichromat", "wesanderson", "Metrics", "car", "DAAG"))
```

```{r echo=FALSE, warning=FALSE}
########################################
### Loading Data
########################################

train <- read.csv('../data/house_price_train.csv')
test <- read.csv('../data/house_price_test.csv')
```

***

### Check for the structure of the dataset
```{r echo=TRUE, warning=FALSE}
glimpse(train)
head(train)
```

<p class="todo">
TODO in FE: Drop `id`
</p>

```{r echo=FALSE, warning=FALSE}
########################################
### DATA CLEANING
########################################
numeric_vars <- unlist(lapply(train, is.integer)) 

for (i in c('id')){
  train[[i]] = as.factor(train[[i]])
}
```

```{r echo=FALSE, warning=FALSE}
########################################
### VARIABLES METADATA
########################################
updateMetaData<- function(ignore, target) {
  TARGET <<- target
  IGNORE <<- ignore
  
  CATEGORICAL <<- unlist(lapply(train, is.factor)) 
  NUMERIC <<- unlist(lapply(train, is.numeric)) 
  
  NUMERIC[IGNORE] <<- FALSE
  CATEGORICAL[IGNORE] <<- FALSE
  NUMERIC[TARGET] <<- FALSE
  CATEGORICAL[TARGET] <<- FALSE
}

updateMetaData(ignore=c("id", "date"), target="price")
```



### Check if there are any missing values in training and test datasets

```{r echo=FALSE, warning=FALSE}
train_na_cols <- colnames(train)[colSums(is.na(train)) > 0]
test_na_cols <- colnames(test)[colSums(is.na(test)) > 0]

print("Train NAs:")
length(train_na_cols)
print("Test NAs:")
length(test_na_cols)
```

**No NAs found**

***

#### Histograms of numerical Features 

```{r echo=FALSE, fig.height=7, warning=FALSE}
histograms = list()

for (i in colnames(train[NUMERIC])){
  histograms[[i]] <-
    ggplot(
      train[NUMERIC],
      aes_string(x=i)) +
      geom_histogram(bins = 40, fill="blue") +
      theme_classic()
}
plot_grid(plotlist = histograms, ncol = 3)
```

**Some skewed distributions**

**Skewed vairables can be a problem for linear regression**

***

#### Histograms against target

**Color coded by Price Quantile 1=lowest 25%, 4= highest 25%**

```{r echo=FALSE, fig.height=10, warning=FALSE}
histograms_against_target = list()

Quantile <- ntile(train$price, 4)

for (i in colnames(train[NUMERIC])){

  histograms_against_target[[i]] <-
    ggplot(train, aes_string(x = i, fill = as.factor(Quantile))) +
    geom_histogram(bins = 40) +
    scale_fill_discrete(name="Price")+
    theme(legend.position = "bottom")+
    theme_classic()
}

plot_grid(plotlist = histograms_against_target, ncol = 2)
```

**Some clear patterns are appear here, for example houses in the highest price quantile (most expensive) only exist at certain latitudes (between 47.5 and 47.7).**

**Pay attention to it in variable correlation analysis!!**

***

#### Variable skewness sorted

```{r echo=FALSE, fig.height=3, warning=FALSE}
skewness <- melt(NUMERIC[NUMERIC==TRUE], value.name = "value")
skewness$var <- rownames(skewness)
                         
for (i in colnames(train[NUMERIC])){
  skewness[i,"value"] = skewness(train[i])
}

skewness <- skewness[order(skewness$value),]

ggplot(data=skewness, aes(y=value, x=reorder(var, value))) +
  geom_bar(stat = "identity", fill="blue") +
  theme_classic()+
  xlab("Variable")+
  ylab("Skewness")+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle("Variables Skewness")
```

***

#### Distribution of Target variable

```{r echo=FALSE, fig.height=4, warning=FALSE}
scaleFUN <- function(x) sprintf("%.0f", x)

pal <- wes_palette("Zissou1", 100, type = "continuous")


g1 <- ggplot(train) + 
        geom_density(aes(x=price), alpha = 0.7)+
        ggtitle("Histogram of Price") +
        scale_x_continuous(labels=scaleFUN)+
        theme_classic()


g2 <- ggplot(train, aes(x = price, fill = ..count..)) +
        geom_histogram(bins=40) +
        scale_fill_gradientn(colours = pal) +
        ylab("Count of houses") +
        xlab("Housing Price") + 
        scale_x_continuous(labels=scaleFUN)+
        theme(plot.title = element_text(hjust = 0.5))

plot_grid(plotlist = list(g1, g2), nrow = 2)

#Summary
summary(train$price)
```

**From these shapes, the distribution of our target variable `price` is skewed to right.**

**One appriach to fix this skewness is taking the log of price.**

***

#### Try taking log of price

```{r echo=FALSE,fig.height=4, warning=FALSE}
#log of price
train$logPrice <- log(train$price)

scaleFUN <- function(x) sprintf("%.0f", x)

g1 <- ggplot(train) + 
        geom_density(aes(x=logPrice), alpha = 0.7)+
        ggtitle("Histogram of LOG Price") +
        scale_x_continuous(labels=scaleFUN)+
        theme_classic()


g2 <- ggplot(train, aes(x = logPrice, fill = ..count..)) +
        geom_histogram(bins=40) +
        scale_fill_gradientn(colours = pal) +
        ylab("Count of houses") +
        xlab("LOG Housing Price") + 
        scale_x_continuous(labels=scaleFUN)+
        theme(plot.title = element_text(hjust = 0.5))

plot_grid(plotlist = list(g1, g2), nrow = 2)

#Summary
summary(train$logPrice)
```

**logPrice is now normally distributed. We can use this with linear models**

<p class="todo">
TODO in FE: log Target Variable (Price)
</p>


***

#### Check date varibale format

```{r echo=TRUE, warning=FALSE}
head(train$date)
```

**It's better to convert it to just number so we can check corrleation with target**

```{r echo=FALSE, warning=FALSE}
# Convert
train$date <- as.Date(train$date, format = "%m/%d/%y")
train$date <- as.integer( format(train$date, "%Y%m%d") )
   
updateMetaData(ignore=c("id"), target="price")
```

***

#### Correlation between two numerical features

```{r echo=FALSE,fig.height=5, warning=FALSE}

correlations <- cor(train[NUMERIC])
col<- colorRampPalette(c("blue", "white", "red"))(20)

corrplot.mixed(
          correlations,
          upper="circle",
          lower="number",
          tl.col = "black",
          tl.pos="lt",
          insig = "blank",
          order="FPC",
          number.cex = .8,
          tl.cex=1)

## Heatmap for correlations
heatmap(x = correlations, col = col, symm = TRUE)
```

**There is a high correlation between some of the variables, this can be a problem with linear regression**

***

# Correlations with Target variable

```{r echo=FALSE,fig.height=3, warning=FALSE}
targetCorr <- data.frame(cor(train[NUMERIC], train$price))
colnames(targetCorr)[1] <- "corr"
targetCorr$var <- rownames(targetCorr)

pal <- wes_palette("Zissou1", 100, type = "continuous")

ggplot(data=targetCorr, aes( y=corr, x=reorder(var, corr))) +
  geom_bar(stat = "identity", aes(fill=corr)) +
  scale_fill_gradientn(colours = pal) +
  theme_classic()+
  xlab("Variable")+
  ylab("p-value")+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle("Correlations with target (price)")

targetCorr
```

<p class="subtitle"> Highly correlated variables with price: </p>

- price and sqft_living have a correlation of 0.7
- price and grade have a correlation of 0.6
- price and sqft_above have a correlation of 0.6
- price and sqft_living15 have a correlation of 0.58
- price and bathrooms have a correlation of 0.53

<p class="subtitle"> Very weakly correlated variables: </p>

- sqft_lot, sqft_lot15, date, yr_built, long

<p class="subtitle"> Needs more analysis: </p>

- condition, waterfront, bathrooms, and floors may be considered as categorical vars so leave them for now.

<p class="todo">
TODO in FE: Drop sqft_lot, sqft_lot15, date, yr_built, long
</p>

```{r echo=FALSE, warning=FALSE}
train$sqft_lot = NULL
train$sqft_lot15 = NULL
train$date = NULL
train$yr_built = NULL
train$long = NULL
```

***

#### Zeroing in on Multicollinearity (most correlated varibales with each other)

```{r echo=FALSE,fig.height=5, warning=FALSE}
signifcentCorrelations <- correlations
signifcentCorrelations[abs(signifcentCorrelations)<0.7] <- 0

corrplot.mixed(
          signifcentCorrelations,
          upper="circle",
          lower="number",
          tl.col = "black",
          tl.pos="lt",
          insig = "blank",
          order="FPC",
          number.cex = .8,
          tl.cex=1)

## Heatmap for correlations
heatmap(x = signifcentCorrelations, col = col, symm = TRUE)
```

<p class="subtitle"> Highly correlated variables with each other: </p>

- sqft_living and sqft_above have a correlation of 0.88
- sqft_above and grade have a correlation of 0.76 (ignore as sqft_above will be drop anyway)
- sqft_living and sqft_living15 have a correlation of 0.76
- sqft_living and grade have a correlation of 0.76
- sqft living and bathrooms have a correlation of 0.76
- sqft_lot and sqft_lot15 have a correlation of 0.72 (ignore as both will be drop anyway)

***

#### Check Multicollinearity between sqft_living and sqft_above have (0.88) using VIF 

**http://www.sthda.com/english/articles/39-regression-model-diagnostics/160-multicollinearity-essentials-and-vif-in-r/**

```{r echo=TRUE, warning=FALSE}
model = lm(formula = price ~ sqft_living + sqft_above,
               data = train)
vif(model)
```

**Very close to 5 (upper limit according to articles)**

**However sqft_living had the higheist collinearity with price.**

<p class="todo">
TODO in FE: Drop sqft_above
</p>

***

#### Check Multicollinearity between sqft_living and sqft_living15 have (0.76) using VIF 

```{r echo=TRUE, warning=FALSE}
model = lm(formula = price ~ sqft_living + sqft_living15,
               data = train)
vif(model)
```

**Below 5, keep both**

***

#### Check Multicollinearity between sqft_living and grade have (0.76) using VIF 

```{r echo=TRUE, warning=FALSE}
model = lm(formula = price ~ sqft_living + grade,
               data = train)
vif(model)
```

**Below 5, keep both**

***

#### Check Multicollinearity between sqft_living and bathrooms have (0.76) using VIF 

```{r echo=TRUE, warning=FALSE}
model = lm(formula = price ~ sqft_living + bathrooms,
               data = train)
vif(model)
```

**Below 5, keep both**

***

#### Potential factor variables

```{r echo=FALSE, fig.height=5, warning=FALSE}

g1 <- ggplot(train, aes(x=as.factor(bedrooms), y=price))+
      geom_boxplot(notch=TRUE)+
      scale_y_continuous(labels=scaleFUN)+
      theme_light()

g2<- ggplot(train, aes(x=as.factor(bathrooms), y=price))+
      geom_boxplot(notch=TRUE)+
      scale_y_continuous(labels=scaleFUN)+
      theme_light()

g3<- ggplot(train, aes(x=as.factor(sqft_living), y=price))+
      geom_boxplot(notch=TRUE)+
      scale_y_continuous(labels=scaleFUN)+
      theme_light()

g4<- ggplot(train, aes(x=as.factor(floors), y=price))+
      geom_boxplot(notch=TRUE)+
      scale_y_continuous(labels=scaleFUN)+
      theme_light()


g5<- ggplot(train, aes(x=as.factor(waterfront), y=price))+
      geom_boxplot(notch=TRUE)+
      scale_y_continuous(labels=scaleFUN)+
      theme_light()

g6<- ggplot(train, aes(x=as.factor(view), y=price))+
      geom_boxplot(notch=TRUE)+
      scale_y_continuous(labels=scaleFUN)+
      theme_light()

g7<- ggplot(train, aes(x=as.factor(condition), y=price))+
      geom_boxplot(notch=TRUE)+
      scale_y_continuous(labels=scaleFUN)+
      theme_light()


g8<- ggplot(train, aes(x=as.factor(grade), y=price))+
      geom_boxplot(notch=TRUE)+
      scale_y_continuous(labels=scaleFUN)+
      theme_light()


plot_grid(plotlist = list(g1, g2, g3, g4), nrow = 2, ncol=2)

plot_grid(plotlist = list(g5, g6, g7, g8), nrow = 2, ncol=2)
```

**bedrooms: is not linearly correlated with target => convert to factor**

**bathrooms: linearly correlated with target => leave it numerical**

**sqft_living: linearly correlated with target => leave it numerical**

**floors: is not linearly correlated with target => convert to factor**

**waterfronts: is not linearly correlated with target => convert to factor**

**view: is not linearly correlated with target => convert to factor**

**condition: is not linearly correlated with target => convert to factor**

**grade: linearly correlated with target => leave it numerical**


**Also bedrooms 11 and 13 seems to be outliers => remove**

<p class="todo">
TODO in FE: Convert bedrooms, floors, waterfronts, view, and condition  to factor
</p>

<p class="todo">
TODO in FE: Remove rows for bedroooms = 11 or 33
</p>

***

